Energy and exergy analysis of a cruise ship

### 2018-05-10 Manuscript/Manuscript.tex
Updates to the manuscript

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm}
%% Tthe amsmath package provides the "align" environment for nicer equation formatting
\usepackage{amsmath}
%% The package tabls helps in having nicer tables with higher spaces
\usepackage{tabls}
%% The package multirow allows merging together multiple cells vertifcally
\usepackage{multirow}
%% The package subcaption is for helping when using subfigures and their captions
\usepackage{subcaption}
%% This package is used for having nicer tables
\usepackage{booktabs}

@@

\subsection{Data processing: Ship energy system modeling} \label{sec:met:processing}

@@

The value of $\dot{Q}_{cooling}$ is determined based on the balance in Equation \ref{eqn:engineEnergyBalance}. The share between the four contributions in Equation \ref{eq:eb:cooling} is calculated based on the data for the heat loss to the HT and LT systems presented in the engines project guides as functions of the engine load. The values are interpolated based on a 2$^nd$ degree polynomial and scaled to respect the energy balance. In absence of more accurate data, it is then assumed that the contributions of $\dot{Q}_{JWC}$ and $\dot{Q}_{LOC}$ are equal for all loads. CHANGE VALUES IN THE TABLE

Values for temperatures, mass flows and energy flows for different engine loads are provided in the Appendix \ref{sec:appendix:engines}, for both main and auxiliary engines.

@@

All values not directly measured in the cooling systems were calculated based on mass and energy balances. The mass flow was determined based on measurements of the pressure in the system and on the performance curves of the cooling pumps.

@@

The total electric power demand of the system is easily estimated as the sum of the power generated by the four auxiliary engines, that is directly measured on the electrical side of the generators and has a rather high accuracy
\begin{equation}
P_{el,tot} = \sum_{1}^{4} AE_i
\end{equation}

The determination of how different consumers contribute to the total electric power demand is less trivial. The ship is equipped with a number of different systems, including lighting, navigational systems, pumps and compressors, etc. None of the individual contributions is directly measured, hence estimations need to be based on indirect measurements.

@@

For what concerns the air conditioning unit demand (HVAC), the observation of the total power demand over the year (see Figure \ref{fig:pelvstime}, showing daily averages) leads to the identification of a clear period, corresponding to the summer months, during which there is a significant increase of the demand.

@@

Since the total demand seems to be rather constant otherwise, it can be concluded that the demand of the HVAC compressors related to the need for cooling is concentrated during the summer months. Additionally, comparing the evolution of the daily demand (see Figure \ref{fig:pelrelvstime}, representing the instantaneous demand divided by the daily average) between a random summer and winter day, shows that the daily variation is comparable. This shows that assuming a constant consumption for the HVAC during the day does not introduce a substantial error in the estimations.

@@

When entering port areas, the ship needs to use thrusters to maneuver and berth. In the case of this particular ships, operating on daily schedules and hence maneuvering four times per day, the power demand related to thrusters can be significant. In order to isolate the demand from the thrusters, we created a reference daily energy profile, based on the instantaneous electric power demand divided by the daily average. From this daily profile (made of 96 points), we selected manually selected the points that could be clearly identified as related to the thrusters energy demand, and substituted the actual value with a weighted average of the previous and subsequent points (see Figure \ref{fig:thruster_selection}).

@@

As the heat demand is not measured, it is necessary to determine it based on the available indirect measurements.

The heat demand and generation can be summarized according to the following equations:
\begin{eqnarray}
\dot{Q}_{gen} & = & \dot{Q}_{EGB} + \dot{Q}_{HTHR} + \dot{Q}_{AB} \\
\dot{Q}_{dem} & = & \dot{Q}_{HVAC,PH} + \dot{Q}_{HVAC,RH} + \dot{Q}_{HWH} + \dot{Q}_{TH} + \dot{Q}_{G} + \dot{Q}_{OT} + 	\dot{Q}_{HTH} + \dot{Q}_{MSH} \\
\end{eqnarray}

The main view of the ship heating and cooling systems is presented in Figure \ref{fig:ship:general}

@@

As not enough information and measurements are available to determine the various components of the heat balance, in this paper we determined them by means of a parameter estimation procedure, using the daily boiler fuel consumption for the calibration of the parameters. The parameter estimation problem is hence written as a minimization problem:
\begin{eqnarray}
min &  \left(\frac{\sum_i(y(\textbf{p})-\bar{y})^2}{\sum_i \bar{y}^2}\right)^{0.5} \\
\end{eqnarray}

where the vector $\textbf{p}$ includes the calibration parameters that are part of the heat demand and generation estimation model that is explained in detail in the following sections. A list of the parameters $\textbf{p}$ is shown in Table REF, together with the chosen upper and lower boundaries for the calibration procedure.
\begin{table}
	\centering
	\begin{tabular}{p{3cm}ccp{1.6cm}p{1.6cm}p{1.2cm}}
		\hline 
		Parameter name & Symbol  & Unit & Lower Boundary & Higher Boundary & Optimal value \\ 
		\hline
		Constant HTHR heat demand	 & $\dot{Q}_{k,HTHR}$ & kW & 0 & 1000 &  \\ 
		Constant steam demand		 & $\dot{Q}_{k,steam}$ & kW & 0 & 1000 &  \\
		Weight factor of the HVAC Re-heater & $f_{HVAC,RH}$ & - & 0.5 & 1 &  \\ 
		Weight factor of the HVAC Pre-heater & $f_{HVAC,PH}$ & - & 0 & 1 &  \\ 
		Weight factor of hot water heater & $f_{HWH}$ & - & 0.5 & 1 &  \\ 
		Weight factor of the galley & $f_{G}$ & - & 0.5 & 1 &  \\ 
		Weight factor of the other consumers & $f_{Other}$ & - & 0.5 & 1 &  \\ 
		HTHR inlet temperature & $T_{HTHR,ER1,in}$ & K & 343 & 353 &  \\
		Effectiveness of the HTHR HEX & $\epsilon_{HTHR} $ & - & 0.5 & 0.9 &  \\
		Boiler drum steam storage capacity & $Q_{ab,max}$ & MJ & 100 & 100000 &  \\ 
		Boiler heat rate & $\dot{Q}_{ab,des}$ & kW & 2000 & 8000 &  \\ 
		\hline
	\end{tabular}
	\caption{Parameters optimized in the parameter estimation for the heat balance}
	\label{tab:ParameterEstimation} 
\end{table}

@@

We calculated the heat demand as the sum of the contributions of the elements listed in the ship's heat balance documentation. As no direct measurement of these quantities was available in the dataset, they had to be estimated based on the following assumptions:
\begin{table}
	\centering
	{\tablinesep=2ex\tabcolsep=10pt
		\begin{tabular}{p{2.8cm}l}
			\hline 
			Heat flow name & Equation \\
			\hline
			HVAC Preheater & $\dot{Q}_{HVAC,PH} = f_{HVAC,RH} \dot{Q}_{HVAC,PH,des} \dfrac{\dot{W}_{HVAC}(t)}{\dot{W}_{HVAC,max}} $ \\
			HVAC Reheater &	$\dot{Q}_{HVAC,RH} = f_{HVAC,PH} \dot{Q}_{HVAC,PH,des} \dfrac{T_{in} - T_{air,out}(t)}{T_{in} - T_{air,out,des}}$ \\
			Hot water heater& $\dot{Q}_{HWH} = f_{HWH} \dot{Q}_{HWH,des} \Phi_{HWH}(\hat{t})$ \\
			Galley & $\dot{Q}_{G} = f_{G} \dot{Q}_{G,des} \Phi_G(\hat{t})$ \\
			Low temperature tank heating & $\dot{Q}_{TH} = f_{TH} \dot{Q}_{TH} \dfrac{T_{T} - T_{air,out}(t)}{T_{T} -T_{air,out,des}}$ \\
			HFO tank heating & $\dot{Q}_{HTH} = f_{HTH} \dot{Q}_{HTH} \dfrac{T_{HT} - T_{air,out}(t)}{T_{HT} - T_{air,out,des}}$ \\
			Machinery space heating & $\dot{Q}_{MSH} = f_{MSH} \dot{Q}_{MSH} \dfrac{T_{MS} - T_{air,out}(t)}{T_{MS} - T_{air,out,des}}$ \\
			HFO heater & $\dot{Q}_{HH} = \dot{m}_{HFO}(t) c_{p,HFO} (T_{HFO,inj} - T_{HT})$ \\
			\hline
	\end{tabular}}
	\caption{Summary of the heat demand contributions and their calculation}
	\label{tab:HeatDemand}
\end{table}

where all $f_i$ factors are treated as calibration parameters (see table \ref{tab:ParameterEstimation}). The $\Phi_G(\hat{t})$ and $\Phi_{HWH}(\hat{t})$ functions represent the assumption made on the daily evolution of the heating demand from the galley and the hot water heater respectively. The daily evolutions of the demand are considered to be the same over the whole year of operations and are represented graphically in Figure REF.

@@

where $T_{eg,EGB,out}$ and $T_{eg,EGB,in}$ are measured for all EGBs, $c_{p,eg}$ is calculated as a function of the exhaust gas composition and temperature, and $ \dot{m}_{eg} $ is calculated based on the engine energy and mass balance as described in section REF and in the appendix REF

It is known that the ship heating systems are designed for recovering energy from the high temperature cooling systems of all the ship's engines. However, measurements of this contribution and of other variables that could lead to its straight-forward identification are missing. In this work, we calculated the heat exchanged in the two HTHRs according to equation \label{eqn:HTHR2},

@@

The results of the parameter estimation are represented graphically and quantitatively in Figure REF and Table REF.

Given the lack of input values for the estimation of the heat demand, the provided estimate based on the procedure described above should be integrated with an estimation of the uncertainty.

@@

The uncertainty on the heat generated in the EGBs can be further subdivided based on its definition as a composition of the uncertainty of $T_{eg,EGB,in}$, $T_{eg,EGB,out}$, $\dot{m}_{eg}$, and $c_{p,eg}$. In the case of $T_{eg,EGB,in}$ and $T_{eg,EGB,out}$, measured values are used, where the uncertainty is related to the sensors (K-type thermocouples) that can be as high as 4\% [CIT]. The uncertainty on the $\dot{m}_{eg}$ is related to the calculation assumptions as described in Appendix REF, and can be estimated of being up to 10\%. The uncertainty of the $c_{p,eg}$ value, given that both composition and temperature are accounted for, should be within 5\% of the reference value. These assumptions lead to the following estimation of the uncertainty:
\begin{equation}
\frac{\delta \dot{Q}_{gen}}{\dot{Q}_{gen}} = \sqrt{\frac{\delta \dot{m}_{eg}^2}{\dot{m}_{eg}^2} + \frac{\delta c_{p,eg}^2}{c_{p,eg}^2} + \frac{\delta T_{EGB,in}^2 + \delta T_{EGB,out}^2}{(T_{EGB,in} - T_{EGB,out})^2}}
\end{equation}

That once typical values are assigned can be as high as 30\%, where most of the variability is related to the temperature measurements (result obtained for $\pm 20$ K uncertainty. The uncertainty is reduced to 20\% if the measurement uncertainty on the temperature is reduced to  $\pm 20$ K)

@@

Figure \ref{fig:vship_hist} represents the distribution of the ship speed. The ship operates almost for the entire year according to a fix schedule, while there are a couple of periods (particularly during the summer months) when the ship operates at higher speed. In Figure \ref{fig:opMode_pie} it can be also observed that the ship spends a relatively high amount of time in port. These remarks suggest that, in spite of the high installed power of the main engines, the energy demand for propulsion is expected to be relatively low.

Figures \ref{fig:loadME_hist} and \ref{fig:loadME_hist} show the frequency of the load at which each of the engines (main and auxiliary engines, respectively) is operated. In the case of the main engines, it can be observed that they are operated mostly at loads between 30\% and 60\%, despite the optimal load for the engines being located around 85\%. This is a consequence of both the low speed at which the ship is generally operated, and of the fact that the two shaft lines are operated independently, and hence do not allow the use of only one of the four engines at high load.

In the case of the auxiliary engines, a larger difference between how the engines are operated can be observed. AE-3 is generally more often operated, as it is the engine that is run in port with clean fuels. On the other hand, it seems that AE-4 is operated rarely, which might be due to maintenance issues.

@@

ANYTHING ELSE THAT CAN BE INTERESTING TO SHOW IN THE EXPLORATORY DATA ANALYSIS?

@@

The ship energy demand is first subdivided among three main consumer categories: propulsion, electric power, and heat. The evolution of these demands over time during a typical voyage is shown in Figures \ref{fig:Demand_vs_time_W} and \ref{fig:Demand_vs_time_S}, representing a winter and a summer day respectively. It can be seen clearly how, as expected, the heating demand in winter is higher than in summer [2800-4500 kW vs 1100-2900 kW], as a consequence of the reduced need for compartment heating. On the other hand, the electric power demand behaves inversely, ranging around 1900 kW during the reference winter day (peaks are connected to the use of thrusters in port for maneuvering) and around 2300 kW during the reference summer day, where the difference is mostly associated to the demand of the HVAC compressors. Looking at the yearly cumulated demand, nearly 50\% of the total energy demand is related to ship propulsion (see Figure \ref{fig:Pie_EnergyDemand}), while the remaining portion is approximately equally split between electric energy and heat demand.

@@

Heat recovery has a large impact on the overall heat demand (see Fig.  \ref{fig:Pie_HeatGeneration}). The the exhaust gas boilers and the heat recovery on the HT-water fulfill almost 75\% of the yearly demand for heating, leaving only the remaining 25\% to be provided by oil-fired auxiliary boilers. The contribution of the HRSG and the HTHR is substantially constant throughout the year (Fig. \ref{fig:TimeSeries_HeatGeneration}), while the auxiliary boilers are used to cover the peaks and the winter increased baseload.

@@

Looking at heating from the demand side, it can be seen that the HVAC pre-heater represents the largest contributor (56\%) to the annual heating demand, followed by the hot water heater (18\%) and by the galley (11\%).In multi-season central HVAC systems the pre-heater is generally used for the main heating contribution during winter, while the re-heater is only used for peak demand, or for adjusting the delivery air temperature during summer, thus explaining the large contribution of the former. The relatively large share of the hot water heater can instead be related to the fact that these systems are used during all seasons.

@@

The results of the exergy analysis of the ship energy systems are shown in Figure REF and Tables REF to REF. 
\begin{itemize}
	\item Table(s) --> Exergy efficiencies
	\begin{itemize}
		\item General at system level
		\item One per system (MEs, AEs, CoolingSystems, SteamSystems) 
	\end{itemize}
	\item Figure --> Grassman diagram
\end{itemize}

@@

It can be observed that there are large irreversibilities concentrated in some specific parts of the system that show potential for improvement. This is particularly the case of the LT/HT merging valves, where exergy is destroyed in the mixing process. Another substantial source of exergy destruction is the steam heater in the HTHR system, where heat at medium pressure (6 bar) is used to heat up water at 90$^o$C. The analysis of exergy destruction also points out the HVAC reheater as a potential source of improvement, based on the use of relatively high-temperature water for heating up air to around 30$^o$C and on the large energy demand of the component. Other important exergy destruction sources can be located in the engines, as expected, but the improvement of engine processes and technology goes beyond the scope of this work. 
\begin{table}
	\centering
	\begin{tabular}{cc}
		\hline 
		Component name & Contribution to exergy losses \\ 
		\hline 
		LT-HT merge & 13.4\% \\ 
		\hline 
		Sea water cooler & 10.6\% \\ 
		\hline 
		Steam heater & 6.0\% \\ 
		\hline 
		HVAC preheater & 6.0\% \\ 
		\hline 
	\end{tabular} 
	\caption{Major contributors to ship exergy losses. Losses related to combustion are not included in the calculation}
	\label{tab:exergyDestruction}
\end{table}

@@

Add reduction of non-HVAC el demand. Say absorption chillers are useless

@@

\section{Engine modelling, figures and tables}

@@

\section{Full tables of energy and exergy flows}

\subsection{Exergy flows}

\begin{table}
	\centering
	\begin{tabular}{clclcrrrrrrrr}
\toprule
\multicolumn{2}{c}{From} & \multicolumn{2}{c}{To}  & Type & Total & Port/Stay & Low Speed Sailing & Maneuvering & High Speed Sailing & Winter & Mid-Season & Summer\\
\midrule
E & Env. & ME & Comp. & Ph & 0.16 & 0.000 & 0.141 & 0.013 & 0.010 & 0.102 & 0.048 & 0.014\\
ME & Comp. & ME & CAC-HT & Ph & 13.64 & 0.004 & 9.932 & 1.148 & 2.555 & 4.976 & 5.626 & 3.037\\
ME & CAC-HT & ME & CAC-LT & Ph & 12.20 & 0.004 & 9.008 & 1.023 & 2.164 & 4.476 & 5.079 & 2.643\\
ME & CAC-LT & ME & Cyl. & Ph & 11.04 & 0.004 & 8.144 & 0.932 & 1.960 & 4.019 & 4.619 & 2.401\\
ME & Turb. & ME & Comp. & W & 20.16 & 0.006 & 15.287 & 1.659 & 3.207 & 7.539 & 8.523 & 4.098\\
ME & Cyl. & ME & Turb. & Ph & 55.91 & 0.037 & 43.951 & 4.839 & 7.085 & 22.177 & 23.408 & 10.327\\
ME & Turb. & E & HRSG & Ph & 15.45 & 0.014 & 12.621 & 1.353 & 1.458 & 6.605 & 6.320 & 2.520\\
ME & Turb. & E & Losses & Q & 0.81 & 0.001 & 0.653 & 0.069 & 0.087 & 0.341 & 0.338 & 0.131\\
ME & Turb. & E & Env. & Ph & 14.04 & 0.014 & 11.256 & 1.285 & 1.489 & 5.694 & 5.941 & 2.409\\
ME & Comp. & ME & BP valve & Ph & 17.44 & 0.004 & 13.179 & 1.425 & 2.831 & 6.557 & 7.323 & 3.559\\
ME & BP valve & ME & Turb. & Ph & 3.80 & 0.001 & 3.246 & 0.277 & 0.275 & 1.581 & 1.697 & 0.522\\
ME & CAC-HT & CS & HT Cool. & Q & 1.16 & 0.000 & 0.924 & 0.125 & 0.392 & 0.501 & 0.547 & 0.394\\
ME & CAC-LT & CS & LT Cool. & Q & 0.59 & 0.000 & 0.864 & 0.091 & 0.203 & 0.456 & 0.460 & 0.242\\

>>>>

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm}
%% Tthe amsmath package provides the "align" environment for nicer equation formatting
\usepackage{amsmath}
%% The package tabls helps in having nicer tables with higher spaces
\usepackage{tabls}
%% The package multirow allows merging together multiple cells vertifcally
\usepackage{multirow}
%% The package subcaption is for helping when using subfigures and their captions
\usepackage{subcaption}
%% This package is used for having nicer tables
\usepackage{booktabs}
% Package for improving table handling
\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

@@

\subsection{Ship energy system modeling} \label{sec:met:processing}

@@

The value of $\dot{Q}_{cooling}$ is determined based on the balance in Equation \ref{eqn:engineEnergyBalance}. The share between the four contributions in Equation \ref{eq:eb:cooling} is calculated based on the data for the heat loss to the HT and LT systems presented in the engines project guides as functions of the engine load. The values are interpolated based on a 2$^nd$ degree polynomial and scaled to respect the energy balance. In absence of more accurate data, it is then assumed that the contributions of $\dot{Q}_{JWC}$ and $\dot{Q}_{LOC}$ are equal for all loads.

Values for temperatures, mass flows and energy flows for different engine loads are provided in the \ref{sec:appendix:engines}, for both main and auxiliary engines.

@@

All values not directly measured in the cooling systems were calculated based on mass and energy balances. Mass flows were determined based on measurements of the pressure in the system and on the performance curves of the cooling pumps as provided in the engine project guides.

@@

The total electric power demand of the system is easily estimated as the sum of the power generated by the four auxiliary engines, that is directly measured on the electrical side of the generators and has a rather high accuracy
\begin{equation}
P_{el,tot} = \sum_{1}^{4} P_{AE_i}
\end{equation}

The ship is equipped with a number of different systems, including lighting, navigational systems, pumps and compressors, etc. However, none of the individual contributions is directly measured, hence estimations of the energy demand of individual consumers need to be based on indirect measurements. In this paper we identified methods for the estimation of the contribution of two consumers: the HVAC systems and the bow thrusters.

@@

The observation of the total power demand over the year (see Figure \ref{fig:pelvstime}, showing daily averages) leads to the identification of a  period, corresponding to the summer months, when the electric power demand is significantly higher then during the remaining months. Based on this observation, in this work we assumed that the demand of the HVAC compressors related to the need for space cooling is concentrated during the summer months. Additionally, comparing the evolution of the daily demand (see Figure \ref{fig:pelrelvstime}, representing the instantaneous demand divided by the daily average) between a typical summer and winter day, shows that the daily variation is comparable. This shows that assuming a constant consumption for the HVAC during the day does not introduce a substantial error in the estimations.

@@

@@

When entering port areas, the ship needs to use thrusters to maneuver and berth. In the case of this particular ships, operating on daily schedules and hence maneuvering four times per day, the power demand related to thrusters can be significant. In order to isolate the demand from the thrusters, we extracted from ship operations a reference daily energy profile, based on the instantaneous electric power demand divided by the daily average. From this daily profile (made of 96 points), we manually selected the points that could be clearly identified as related to the thrusters energy demand, and substituted the actual value with a weighted average of the previous and subsequent points (see Figure \ref{fig:thruster_selection}).

@@

The case study ship is not provided with any means to directly measure the heat demand (e.g. steam flow meters), and it was hence necessary to provide estimations based on other measurements and assumptions. The heat demand is subdivided among the contributions of the HVAC preheater and reheater ($\dot{Q}_{HVAC,PH}$ and $\dot{Q}_{HVAC,RH}$), the hot water heater ($\dot{Q}_{HWH}$), the HFO heaters ($\dot{Q}_{HH}$) the tank heating systems for HFO and other liquids ($\dot{Q}_{HTH}$ and $\dot{Q}_{TH}$), the machinery space heating ($\dot{Q}_{MSH}$) and other tanks ($\dot{Q}_{OT}$). The heat generation is instead composed of the contribution of the exhaust gas boilers ($\dot{Q}_{EGB,ME,i}$ and $\dot{Q}_{EGB,AE,i}$), the heat recovery from the high temperature cooling systems ($\dot{Q}_{HTHR,j,i}$) and the boilers ($\dot{Q}_{AB}$).
 The heat demand and generation are summarized in equations \ref{eq:Qgen} and \ref{eq:Qdem}.
\begin{align}
\dot{Q}_{gen} & = \sum_{i=2,3} \dot{Q}_{EGB,ME,i} +  \sum_{i=1:4} \dot{Q}_{EGB,AE,i} + \sum_{j=ME,AE, i=1:4}\dot{Q}_{HTHR,j,i} + \dot{Q}_{AB} \label{eq:Qgen} \\
\dot{Q}_{dem} & = \dot{Q}_{HVAC,PH} + \dot{Q}_{HVAC,RH} + \dot{Q}_{HWH} + \dot{Q}_{HH} + \dot{Q}_{TH} + \dot{Q}_{G} + \dot{Q}_{OT} + 	\dot{Q}_{HTH} + \dot{Q}_{MSH} \label{eq:Qdem} \\
\end{align}

The main view of the ship heating and cooling systems is presented in Figure \ref{fig:ship:general}.

@@

The individual contributions of different consumers were determined by means of a parameter estimation procedure, using the daily boiler fuel consumption for the calibration of the parameters. The parameter estimation problem is hence written as a minimization problem:
\begin{eqnarray}
min &  \left(\frac{\sum_i(y(\textbf{p})-\bar{y})^2}{\sum_i \bar{y}^2}\right)^{0.5} \\
\end{eqnarray}

where the vector $\textbf{p}$ includes the calibration parameters that are part of the heat demand and generation estimation model that is explained in detail in the following sections. A list of the parameters $\textbf{p}$ is shown in Table \ref{tab:ParameterEstimation}, together with the chosen upper and lower boundaries for the calibration procedure.
\begin{table}
	\centering
	\begin{tabular}{p{3cm}ccp{1.6cm}p{1.6cm}p{1.2cm}}
		\hline 
		Parameter name & Symbol  & Unit & Lower Boundary & Higher Boundary & Optimal value \\ 
		\hline
		Constant HTHR heat demand	 & $\dot{Q}_{k,HTHR}$ & kW & 0 & 1000 &  \\ 
		Constant steam demand		 & $\dot{Q}_{k,steam}$ & kW & 0 & 1000 &  \\
		Weight factor of the HVAC Re-heater & $f_{HVAC,RH}$ & - & 0.5 & 1 &  \\ 
		Weight factor of the HVAC Pre-heater & $f_{HVAC,PH}$ & - & 0 & 1 &  \\ 
		Weight factor of hot water heater & $f_{HWH}$ & - & 0.5 & 1 &  \\ 
		Weight factor of the galley & $f_{G}$ & - & 0.5 & 1 &  \\ 
		Weight factor of the other consumers & $f_{Other}$ & - & 0.5 & 1 &  \\ 
		HTHR inlet temperature & $T_{HTHR,ER1,in}$ & K & 343 & 353 &  \\
		Effectiveness of the HTHR HEX & $\epsilon_{HTHR} $ & - & 0.5 & 0.9 &  \\
		Boiler drum steam storage capacity & $Q_{ab,max}$ & MJ & 100 & 100000 &  \\ 
		Boiler heat rate & $\dot{Q}_{ab,des}$ & kW & 2000 & 8000 &  \\ 
		\hline
	\end{tabular}
	\caption{Parameters optimized in the parameter estimation for the heat balance}
	\label{tab:ParameterEstimation} 
\end{table}

@@

The heat demand was calculated as the sum of the contributions of the elements listed in the ship's heat balance documentation. The assumptions used in the estimation are summarized in Table \ref{tab:HeatDemand}:
\begin{table}
	\centering
	{\tablinesep=2ex\tabcolsep=10pt
		\begin{tabular}{p{2.8cm}l}
			\hline 
			Heat flow name & Equation \\
			\hline
			HVAC Preheater & $\dot{Q}_{HVAC,PH} = f_{HVAC,RH} \dot{Q}_{HVAC,PH,des} \dfrac{\dot{W}_{HVAC}(t)}{\dot{W}_{HVAC,max}} $ \\
			HVAC Reheater &	$\dot{Q}_{HVAC,RH} = f_{HVAC,PH} \dot{Q}_{HVAC,PH,des} \dfrac{T_{in} - T_{air,out}(t)}{T_{in} - T_{air,out,des}}$ \\
			Hot water heater& $\dot{Q}_{HWH} = f_{HWH} \dot{Q}_{HWH,des} \Phi_{HWH}(\hat{t})$ \\
			Galley & $\dot{Q}_{G} = f_{G} \dot{Q}_{G,des} \Phi_G(\hat{t})$ \\
			Low temperature tank heating & $\dot{Q}_{TH} = f_{TH} \dot{Q}_{TH} \dfrac{T_{T} - T_{air,out}(t)}{T_{T} -T_{air,out,des}}$ \\
			HFO tank heating & $\dot{Q}_{HTH} = f_{HTH} \dot{Q}_{HTH} \dfrac{T_{HT} - T_{air,out}(t)}{T_{HT} - T_{air,out,des}}$ \\
			Machinery space heating & $\dot{Q}_{MSH} = f_{MSH} \dot{Q}_{MSH} \dfrac{T_{MS} - T_{air,out}(t)}{T_{MS} - T_{air,out,des}}$ \\
			HFO heater & $\dot{Q}_{HH} = \dot{m}_{HFO}(t) c_{p,HFO} (T_{HFO,inj} - T_{HT})$ \\
			\hline
	\end{tabular}}
	\caption{Summary of the heat demand contributions and their calculation}
	\label{tab:HeatDemand}
\end{table}

where all $f_i$ factors are treated as calibration parameters (see table \ref{tab:ParameterEstimation}). The $\Phi_G(\hat{t})$ and $\Phi_{HWH}(\hat{t})$ functions represent the assumption made on the daily evolution of the heating demand from the galley and the hot water heater respectively. The daily evolutions of the demand are considered to be the same over the whole year of operations and are represented graphically in Figure \ref{fig:galleyhotwater}.

\begin{figure}[tbp!]
	\centering
	\includegraphics[width=0.7\linewidth]{Figures/galleyHotWater.pdf}
	\caption{Hourly relative demand of hot water heating and galley}
	\label{fig:galleyhotwater}
\end{figure}

@@

where $T_{eg,EGB,out}$ and $T_{eg,EGB,in}$ are measured for all EGBs, $c_{p,eg}$ is calculated as a function of the exhaust gas composition and temperature, and $ \dot{m}_{eg} $ is calculated based on the engine energy and mass balance.

It is known that the ship heating systems are designed for recovering energy from the high temperature cooling systems of all the ship's engines. Based on the available measurements, we calculated the heat exchanged in the two HTHRs according to equation \label{eqn:HTHR2},

@@

@@

The uncertainty on the heat generated in the EGBs can be further subdivided based on its definition as a composition of the uncertainty of $T_{eg,EGB,in}$, $T_{eg,EGB,out}$, $\dot{m}_{eg}$, and $c_{p,eg}$. In the case of $T_{eg,EGB,in}$ and $T_{eg,EGB,out}$, measured values are used, where the uncertainty is related to the sensors (K-type thermocouples) that can be as high as 4\% [CIT]. The uncertainty on the $\dot{m}_{eg}$ is related to the calculation assumptions and can be estimated of being up to 10\%. The uncertainty of the $c_{p,eg}$ value, given that both composition and temperature are accounted for, should be within 5\% of the reference value. These assumptions lead to the following estimation of the uncertainty:
\begin{equation}
\frac{\delta \dot{Q}_{gen}}{\dot{Q}_{gen}} = \sqrt{\frac{\delta \dot{m}_{eg}^2}{\dot{m}_{eg}^2} + \frac{\delta c_{p,eg}^2}{c_{p,eg}^2} + \frac{\delta T_{EGB,in}^2 + \delta T_{EGB,out}^2}{(T_{EGB,in} - T_{EGB,out})^2}}
\end{equation}

That once typical values are assigned can be as high as 30\%, where most of the variability is related to the temperature measurements.

@@

Figure \ref{fig:vship_hist} represents the distribution of the ship speed. The ship operates for almost  the entire year according to a fix schedule, while there are a couple of periods (particularly during the summer months) when the ship operates at higher speed. In Figure \ref{fig:opMode_pie} it can be also observed that the ship spends a relatively large amount of time in port. These remarks suggest that, in spite of the high installed power of the main engines, the energy demand for propulsion is expected to be relatively low.

Figures \ref{fig:loadME_hist} and \ref{fig:loadAE_hist} show the frequency of the load at which each of the engines (main and auxiliary engines, respectively) is operated. In the case of the main engines, it can be observed that they are operated mostly at loads between 30\% and 60\%, despite the optimal load for the engines being located around 85\%. This is a consequence of both the low speed at which the ship is generally operated, and of the fact that the two shaft lines are operated independently, and hence do not allow the use of only one of the four engines at high load.

In the case of the auxiliary engines, a larger difference between how the engines are operated can be observed. AE-3 is generally more often operated, as it is the engine that is run in port with clean fuels. On the other hand, it seems that AE-4 is operated rarely, which might be due to maintenance issues during the selected period of analysis.

@@

@@

The ship energy demand is first subdivided among  The evolution of the demands of propulsion, electric power, and heat over time during a typical voyage is shown in Figures \ref{fig:Demand_vs_time_W} and \ref{fig:Demand_vs_time_S}, representing a winter and a summer day respectively. As expected, the heating demand in winter is higher than in summer [2800-4500 kW vs 1100-2900 kW], as a consequence of the reduced need for compartment heating. On the other hand, the electric power demand behaves inversely, ranging around 1900 kW during the reference winter day (peaks are connected to the use of thrusters in port for maneuvering) and around 2300 kW during the reference summer day, where the difference is mostly associated to the demand of the HVAC compressors. Looking at the yearly cumulated demand, nearly 50\% of the total energy demand is related to ship propulsion (see Figure \ref{fig:Pie_EnergyDemand}), while the remaining portion is approximately equally split between electric energy and heat demand.

@@

Heat recovery has a large impact on the overall heat demand (see Fig.  \ref{fig:Pie_HeatGeneration}). The exhaust gas boilers and the heat recovery on the HT-water fulfill almost 75\% of the yearly demand for heating, leaving only the remaining 25\% to be provided by oil-fired auxiliary boilers. The contribution of the HRSG and the HTHR is substantially constant throughout the year (Fig. \ref{fig:TimeSeries_HeatGeneration}), while the auxiliary boilers are used to cover the peaks and the winter increased baseload.

@@

On the heating demand side, the HVAC pre-heater represents the largest contributor (56\%) to the annual heating demand, followed by the hot water heater (18\%) and by the galley (11\%). In multi-season central HVAC systems the pre-heater is generally used for the main heating contribution during winter, while the re-heater is only used for peak demand, or for adjusting the delivery air temperature during summer, thus explaining the large contribution of the former. The relatively large share of the hot water heater can instead be related to the fact that these systems are used during all seasons.

@@

The results of the exergy analysis of the ship energy systems are shown in Figure \ref{fig:Grassmann} and in Tables REF to REF in the Appendix. 
\begin{itemize}
	\item Table(s) --> Exergy efficiencies
	\begin{itemize}
		\item General at system level
		\item One per system (MEs, AEs, CoolingSystems, SteamSystems) 
	\end{itemize}
	\item Figure --> Grassman diagram
\end{itemize}

@@

It can be observed that there are large irreversibilities concentrated in some specific parts of the system that show potential for improvement. The individual largest contribution to exergy destruction in the ship systems is associated to combustion, both in the engines and in the boilers. The loss of energy resulting from the conversion of chemical energy to heat contributes to more than 76\% of the total exergy destruction on board. Once this contribution is isolated, the largest sources of exergy destruction are located in the engine turbochargers (28.9\%), in the exhaust gas boilers (10.7\%), in the steam heater of the HTHR system (10.5\%), the HVAC preheater (10.4\%) and the sea water cooler (8.5\%). In all the cases above, excluding that of the turbochargers, the exergy destruction is a consequence of a combination of large exergy flows and of the mismatch between hot and cold streams of the heat exchangers. In the exhaust boilers, exhaust gas are cooled in average from more than 300$^o$C to 200$^o$C, while the heat is used to generate steam at relatively low pressure (6 bar); in the steam heater, the steam is used to heat up water at 90$^o$C; in the HVAC preheater, the relatively high-temperature water from the HTHR systems is used for heating up air to around 30$^o$C.

\begin{table}
	\centering
	\begin{tabular}{cc}
		\hline 
		Component name & Contribution to exergy losses \\ 
		\hline 
		Turbochargers & 28.9\% \\ 
		HRSGs & 10.7\% \\
		Steam heater & 10.5\% \\ 
		HVAC preheater & 10.4\% \\
		Sea water cooler & 8.5\% \\ 
		Electric generators & 6.1\% \\ 
		\hline 
	\end{tabular} 
	\caption{Major contributors to ship exergy losses. Losses related to combustion are not included in the calculation}
	\label{tab:exergyDestruction}
\end{table}

@@

The analysis also suggests that the electric power demand resulting from the use of the HVAC on board is relatively low, a consequence of the fact that the case study ship mostly operates in cold climates. This suggests that demand-related efforts should be focused on other consumers (e.g. lighting) rather than on the HVAC compressor. Also, the analysis suggests that the use of related systems for WHR, such as absorption chillers, would not be particularly beneficial in the specific case here studied.

@@

\section{Engine modelling, figures and tables} \label{sec:appendix:engines}

@@

\clearpage

\section{Full tables of energy and exergy flows} \label{sec:appendix:energy}

\subsection{Energy flows}
